---
title: "Data Analysis Project | Analysing LGBTQ Pop, Unemployment and COVID"
author: "Philip Van Slooten"
date: "5/07/2020"
output:
  html_document:
  toc: true
toc_depth: 3
toc_float: true
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, paged.print=TRUE)
```

# Data Analysis Project: LGBTQ Population, Unemployment & COVID-19

## Data Story Pitch: COVID-19's economic impact on the LGBTQ community -- The Curious Case of Nevada

The restaurant and hospitality industries were among the first impacted during the early phases of the social distancing restrictions implemented by U.S. states during the height of the COVID-19 pandemic throughout the month of March. 

To slow the spread of the virus and "flatten the curve," governors across the country implemented stay-at-home orders and closed many other "nonessential" businesses, but the restaurant and hospitality industry continued to show the highest unemployment levels, according to March 2020 Bureau of Labor Statistics data. 

Nevada, with its hospitality-driven economy, ranked second behind Louisiana, another tourism-driven state, in terms of March unemployment numbers and third behind D.C. and Oregon in terms of LGBTQ population, according to Williams Institute data. 

This makes Nevada an interesting case to analyze in terms of the economic impact of COVID-19 restrictions on the LGBTQ community. 

For this assignment I have chosen to visualize two data sets to compare Nevada against other states in terms of LGBTQ population and COVID-related unemployment: 

* U.S. LGBT population [https://williamsinstitute.law.ucla.edu/visualization/lgbt-stats/?topic=LGBT&area=32#density], 

* U.S. Unemployment for March 2020 [https://www.bls.gov/news.release/laus.t01.htm] and


While the purpose of this visualization is to get a picture of COVID-19's economic impact on the LGBTQ community, it is limited by the lack of data collection on the U.S. LGBTQ population. 


# Visualizing Nevada in terms of LGBTQ population and unemployment levels

U.S. LGBTQ aggregate data is difficult to find since both the Census Bureau and the Bureau of Labor Statistics currently refuse to collect demographic information based on sexual orientation or gender identity. 

However, UCLA's Williams Institute does collect and report national LGBTQ data and found high numbers of LGBTQ in lower-income jobs including in the food and hospitality industries which has been hard hit by COVID-19 restrictions. 

Hypothesis: More LGBTQ, Higher Unemployment

As a result, my hypothesis is that April 2020 BLS data when joined with LGBTQ population data should show as LGBTQ population rises, so should unemployment. 

Preliminary Findings: No visual correlation; Nevada outlier

When I visually compared BLS March 2020 unemployment, LGBTQ population and COVID-19 data I did not see a correlation. However, I did see an outlier that fit my hypothesis: Nevada. 

Nevada ranked 2nd in March unemployment and 3rd in LGBTQ population density (but 28 in COVID-19 cases as of April 2020). 

What is driving the high LGBTQ unemployment in the state? It could be because most were employed in the hospitality industry out there (which BLS confirmed is the major industry employer for the state).


## Data Analysis: The Investigation

# Step 1: The Packages

```{r}
#Install the suite of packages below:

#install.packages('tidyverse')
#install.packages('janitor')
#install.packages('rvest')
#install.packages('corrr')
#install.packages('tidycensus')
#install.packages('scales')
#install.packages('mapview')
#install.packages('ggthemes')
#install.packages ('ggrepel')


#Loaded libraries listed below:

library('tidyverse')
library('janitor')
library('rvest')
library('corrr')
library('tidycensus')
library('scales')
library('mapview')
library('ggthemes')
library('ggrepel')


```
 
# Step 2: The Data

```{r}

# Web Scrape 1: March 2020 BLS Table 1. Civilian labor force and unemployment by state and selected area, seasonally adjusted

unemployment_scrape<- read_html("https://www.bls.gov/news.release/laus.t01.htm") %>% html_nodes('table') %>% html_table(header=1, fill=TRUE) %>% as.data.frame()



# Web Scrape 2: LGBT Proportion of Population - United States 2019 [Williams Institute, UCLA]

LGBTQ_scrape<- read_html("https://docs.google.com/spreadsheets/d/1M0uuK9UqsgS2rQpzVBm1X0eWcwoiXBS0TUjTUr7oXBg/edit?usp=sharing")%>% html_nodes('table') %>% html_table(header=1, fill=TRUE) %>% as.data.frame() 



```
 
# Step 3: Data Cleaning



```{r}

# unemployment_scrape_working = clean BLS March 2020 dataset 

unemployment_scrape_working<- unemployment_scrape %>% clean_names()%>% mutate (state=state_and_area)%>% select(state, unemployed_7, everything())%>% select(-state_and_area)%>% mutate (Mar2020_percent_unemployed=unemployed_7)%>% select(-unemployed_7)%>% select(state, Mar2020_percent_unemployed, everything())%>% filter(!str_detect(state,"State and area"))%>% filter(!str_detect(state, "District of Columbia"))%>% filter(!str_detect(state, "Los Angeles"))%>% filter(!str_detect(state, "Miami")) %>% filter(!str_detect(state,"Chicago")) %>% filter(!str_detect(state, "Detroit")) %>% filter(!str_detect(state, "Cleveland")) %>% filter(!str_detect(state,"Seattle")) %>% filter(!str_detect(state, "Puerto Rico")) %>% filter(!str_detect(state, "Footnotes")) %>% filter(!str_detect(state,"Data"))%>% filter(!str_detect(state,"New York City"))%>% mutate (Mar2020_percent_unemployed=as.numeric(Mar2020_percent_unemployed))


```


```{r}

# LGBTQ_scrape_working = clean Williams Institute 2019 dataset

LGBTQ_scrape_working<-LGBTQ_scrape %>% clean_names() %>% mutate(state=a)%>% mutate (percent_LGBTQ=b)%>% select(state, percent_LGBTQ, everything())%>% select(-a,-b) %>% filter(!str_detect(state,"State"))%>% filter(!str_detect(state, "District of Columbia"))%>% mutate (percent_LGBTQ=as.numeric(percent_LGBTQ))

```


# Step 4: Enterprise Join of LGBTQ Pop and Mar 2020 Unemployment data

```{r}

Mar2020_LGBTQ_unemployment<- unemployment_scrape_working %>% inner_join(LGBTQ_scrape_working, by="state") %>% select(state, Mar2020_percent_unemployed, percent_LGBTQ, everything())

```

# Step 5: Data Visualization (Scatterplot)

```{r}
ggplot(Mar2020_LGBTQ_unemployment) + geom_point(aes(percent_LGBTQ, Mar2020_percent_unemployed)) + labs(x="% LGBTQ", y="% Unemployed", title = "State-level LGBTQ and Unemployment Data for March 2020", caption = "Sources: Williams Institute and Bureau of Labor Statistics", fill="state") + scale_y_continuous(labels=comma) + scale_x_continuous(labels=comma) + theme(axis.text.x=element_text(angle=45,hjust=1)) + geom_smooth(aes(percent_LGBTQ, Mar2020_percent_unemployed), method="lm", se=FALSE) + geom_text_repel(aes(percent_LGBTQ, Mar2020_percent_unemployed, label=state))

```

# Step 6: Strength Test

```{r}
Mar2020_LGBTQ_unemployment %>% ungroup() %>% select(percent_LGBTQ, Mar2020_percent_unemployed) %>% correlate()
# The r= .20 (general relationship = positive // strength = weak). 
```
#Calculate the significance. 

```{r}
correlation_lgbtq_unemployment<-Mar2020_LGBTQ_unemployment %>% ungroup() %>% select(percent_LGBTQ, Mar2020_percent_unemployed)

#Test the relationship.
cor.test(correlation_lgbtq_unemployment$percent_LGBTQ, correlation_lgbtq_unemployment$Mar2020_percent_unemployed)

# p-value around 0.15 which is larger than .05 but within the 95% confidence interval. 

#So, nationally there is a significant but mildly positive relationship between LGBTQ population size and COVID-related unemployment during March 2020. 

#But the real story is with the high results for Nevada. 
```


## Preliminary Conclusions

	I have confirmed Nevada’s position as an outlier in terms of LGBTQ population and COVID-related unemployment for March, using scatterplot and correlation analysis.
	
	General findings indicate a mildly positive and slightly significant correlation between LGBTQ population size and COVID-related unemployment. This means that unemployment was slightly higher in states with higher LGBTQ population. However, this trend was most pronounced in Nevada, which also has a leisure-industry based economy. 
	
  Interviews with a few of Nevada's LGBTQ residents  could make for an interesting feature story or profile illustrating COVID's economic impact on the LGBTQ community.

 
 




